\documentclass{article}
\usepackage{amsmath}
\usepackage{geometry}
\usepackage{algorithm}
\usepackage[noend]{algpseudocode}

\author{Hizbullah Abdul Aziz}
\title{Notes}

\begin{document}
\maketitle

\section{Graph Algorithms}
Graph $G=(V,E)$ is defined with a set of vertices $V$ and set of edges $E$.
Now, some notations:
\begin{description}
  \item[Sizes] $m=|E|$ and $n=|V|$
  \item[Directed edge] $[u,v\rangle=\langle v,u]$ denotes an edge from $u$ to
    $v$
  \item[Out- and In- degree] number of edges going out and in of vertex $v$,
    $Out(v)$ and $In(v)$ respectively
\end{description}

There are two ways to represent graphs: adjacency matrices and adjacency lists.
Adjacency matrix needs space $O(n^2)$, which is inefficient for sparse graphs.
Conversely, adjacency list needs space $$\sum_{u\in V} (Out(v) + 1)=|E|+n=m+n$$
which is efficient for sparse graphs.


\subsection{Searching}
\subsubsection{Depth-First Search}
Depth-First Search creates a search tree by exploring each node until it reach
some node that doesn't have any outgoing nodes or the only outgoing nodes are
already visited.
\begin{algorithm}
  \caption{Depth-First Search}\label{graph:dfs}
  \begin{algorithmic}
    \Function{DFS}{g}
    \ForAll{v in V}\Comment{Initialization}
      \State $color[v]=\texttt{white}$
      \State $\pi[v]=\texttt{nil}$\Comment{Parent pointer}
    \EndFor
    \State $time = 0$
    \ForAll{v in V}
      \If{$color[v]=\texttt{white}$}
        \State\Call{DFS-visit}{v}
      \EndIf
    \EndFor
    \EndFunction
    \State
    \Function{DFS-visit}{u}
      \State $color[u]=\texttt{grey}$
      \State $d[u]=t+1$\Comment{Discovery time}
      \ForAll{v in Out[u]}
      \If{$color[v]=\texttt{white}$}
        \State\Call{DFS-visit}{v}
        \State $\pi[v]=u$
      \EndIf
      \State $color[u]=\texttt{black}$
      \State $f[u]=time+1$\Comment{Finishing time}
      \EndFor
    \EndFunction
  \end{algorithmic}
\end{algorithm}


% This is the start of the flow algorithms
\subsection{Flow Problem}
Given a directed graph $G=(V, E)$, we can construct a flow network by defining
a capacity function $u: E\rightarrow R$ and a flow function $f: E\rightarrow
R$, with relation as follows.

Residual graph
It must be possible to undo the flow

\subsubsection{Ford Fulkerson}
In residual graph, find augmenting path until t is not reachable from
s i.e. there's an edge with residual flow equal 0

\subsubsection{Edmunds-Karp Algorithm}
As augmenting path, always choose an s-t path with smallest number of edges,
which can be found via BFS in $O(m + n)$ time.

with BFS, we can't have down edges (if you put s at top and t somewhere below it) that jumps a level

augmentation along shortest path:
- saturate at least one down edge effectively kills it in the residual graph
- creates only up edges
Note that this cannot shorten any s-t path

Observation: can do saturation at most m times
=> after <= m iterations distance between s and t must increase
Distance can increase <n times
This means total number of augmentations is less than $mn$
Since finding an augmenting path takes linear time, then total running time is $O(m^2n)$

\subsubsection{Blocking flows algorithm}
Level network (BFS on graph)
f is blocking flow in level network if $G_f$ admits no downward s-t path

repeat
	form level graph in $G_f$ (BFS starting from s)
	find a blocking flow $f'$ in $G_f$
	$f = f + f'$
until s and t are not connected in $G_f$

Note that like EK alg above, each iteration increases distance between s and
t in $G_f$, so there are at most <n iterations.

Now how to find blocking flow?
DFS to find s-t path
	assume no backtrack -> augment this path, then saturates one edge

	if you backtrack along an edge (t cannot be reached along this edge)
		you can remove this edge from consideration

throughput(v) = min{capacity out, capacity in}

Consider case where all edges have capacity 1.  Then finding a blocking flow takes O(m) time


\subsubsection{Unit Network}
All edges have capacity 1.
For all nodes v in V\{s,t}, either indeg(v) = 1 or outdeg(v) = 1

Observation: any residual graph of a unit network is also a unit network.  
Why?  Take any node and see that the resulting residual graph is basically the
flow network but with the edges reversed.

Claim: Given unit network, and shortest path from s to t has $k+1$ edges.  This
implies the max flow can have value at most $n/_k$
Proof: Need to show s-t cut with capacity at most $n/_k$
Make a level graph.  Then there's a level with at most $n/_k$ vertices.  Then
take any vertices with one node coming in as the t part and any vertices with
one node coming out as the s part, forming a cut.  Since this level has at most
$n/_k$ vertices, then the cut has $n/_k$ edges coming in and out.  Since every
edge has capacity 1, then $n/_k$ is the capacity of the cut, which is also the
max flow.

Algorithm for unit network (Karp, Hopcroft)
1. Run $\sqrt n$ iterations of blocking flow
2. Run $\sqrt n$ augmenting path finding


\subsection{Circulation Problem}
Network with edge capacities c: E -> Real but no source and sink nodes
But we have supply/demand b: B -> Real (positive for supply, negative for
demand)

Constraint: b(V) = 0, that is, sum of all supply at all vertices is zero

Want: flow func f: E -> positive Real such that for every vertices v,
	Inflow(v) - Outflow(v) + b(v) = 0

Solutions need not exists!
Therefore the question really is, is the problem feasible?  That is, is there
such f?

It turns out we can solve this using the max flow
Lemma: for every circulation problem in G= (V, E; c, b), there exists a max flow
problem G'= (V', E'; c') s, t such that circulation problem is feasible iff max
flow value for G' is the sum of all supplies.


\subsubsection{Min Cost Circulation Problem}
Cycle cancelling
1. Start with a feasible circulation e.g. $\forall a f(a) = 0$
2. While there's a negative directed circuit $Z\subseteq A^f$ in $G^f$
   a. compute $\tau = \min \{u^f(a) \in Z\}$
   b. update every edge $a$: $f(a) = f(a) + \tau if a \in Z f(a) - \tau if -a \in Z f(a) otherwise$
   c. return f

Theorem: A feasible circulation $f: A -> Z$ is of minimum cost iff
$G^f$ does not contain a negative directed circuit.
Proof:
1. Necessity: suppose there's a negative directed circuit. Thus we
could update and get a bigger flow, and thus even smaller cost.
2. Sufficiency: Suppose that each directed circuit in $G^f$ has non-negative cost.  Now we can show that f is indeed the min cost

Theorem: Choosing a Minimum Mean Cycle (MMC) in each iteration leads to at most $4nm^2\ceil \ln n$
Proof:
Let $f, f^{-1},\ldots$ be the circulation found in.  For each 

\subsubsection{Reduction to regular circulation problem}
do some weird shit

% This is the end of the flow algorithms


\subsection{Matching}
$G=(V,E)$ undirected graph
Match $M\in E$ such that every vertices incident to at most 1 edge in M

M is called maximal iff there's no matching M' with M strictly inside M.  In
other words, can't add more edges to M.

M is called maximum if |M| is as large as possible.

\subsubsection{Matching in Bipartite Graph}
G is bipartite if we can partition V into A and B and all edges are between
vertices in A and vertices in B.

How to find the maximum matching?  Use the unit network algorithm in time
O(m\sqrt n)

\subsubsection{Matching in General Graph}
Alternating path in a graph with matching M is a path whose edges alternate
between in M and not in M.

Augmenting alternating path if it starts and ends in two distinct unmatched
vertices.

Augmenting path lemma:
Graph $G=(V,E)$ and matching $M\in E$.  M is maximum iff there's no augmenting
alternating path.
Proof: 
- if there's an augmenting alternating path then we can always construct a bigger matching.
- assume M is not maximum and M* be the maximum matching.
	1 could be a cycle (same number of edge for M and M*)
	2 could be a path
	3 could be that M and M* contain the same edges
  since M* is bigger, then there must be situation 2, where we get the augmenting alternating path.

Notion of stem, root, and blossom (which always with respect to some matching M)
Blossom is basically an odd cycle attached to root.
Stem always starts with a free vertex and has even edges.

Shrinking a blossom -> contract blossom to a single vertex.

Theorem:
Graph $G=(V,E)$, matching M and B blossom with respect to M.  There exists an
augmenting alternating path in G iff there exists an augmenting alternating
path in G with B shrinked.

Alg to find an augmenting alternating path in O(mn)
M matching
free vertices
matched vertices w, mate(w) -> M = {w,mate(w)}
Find "alternating" forest
	node are marked with "unlabelled" "odd" "even"
Initially all free vertices are even and matched vertices are unlabelled.

Pick some even node v and let w some adjacent vertex
Explore edge {v,w}.  Now there are several cases
1. w is unlabelled -> make w child of v; make mate(w) child of w; label w as odd; label mate(w) as even
2. w is odd -> do nothing
3. w is even -> if in different tree then found an alternating path
		else found a blossom; shrink it and start anew

Therefore we can find a matching in $O(n^2m)$


\end{document}
